Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-06-01 manuscript.tex
Bring .tex current

\abstract{
Social mobilization requires effective collective coordination. In recent
years, individuals involved in social movements have taken to social media
platforms to reach broader audiences at rapid speeds. Effective virtual
mobilization similarly requires social cohesion and coordination, yet the
degree to which virtual mobilization reflects or drives real-world movements
is unclear. Here, we explore the dynamics of real-world events and social
cohesion on Twitter in Syria during the Arab Spring. Using the nonlinear
methods cross-recurrence quantification analysis and windowed cross-recurrence
quantification analysis, we investigate if counts of all daily events, daily
positive events, and daily negative events are coupled with social cohesion
found in Syrian tweets. Results indicate that virtual social cohesion is
coupled with the counts of all and negative events each day but not with
counts of positive events. This suggests that virtual mobilization does reflect
real-world movements but only for salient, frequent and negative events. We
conclude with a discussion of implications and applications of nonlinear
methods in political science research.
}

@@

Social coordination has long been used both physically and virtually
to start and carry out collective actions. These collective events
bring individuals together and create emergent patterns of communication
and action. Online communication has historically been used to document
and communicate archival information about events, but it has been
increasingly used to bring about social change. From a dynamical
systems perspective, these different means of social coordination during
times of turmoil can be considered two complementary components of a
single system, mutually constrained and constantly interacting and
influencing one another (Mahner \& Bunge, 1997). Investigation of the
coupling of real-world mobilization and social cohesion could provide
new insight into the power of social media during mobilization and the
ways in which individuals utilize it to bring about, or respond to, social
change. Increasingly, social media platforms are being utilized during
social mobilization, but the relationship between virtual and real-world
mobilization is still largely unknown. The present work aims to understand
how physical and virtual movements coexist and interact from a dynamical
systems perspective.

\hypertarget{arab-spring}{%
\subsection{Arab Spring}\label{arab-spring}}

@@

These mass movements would not have been possible without
efficient large-scale coordination, both physically on the ground and virtually
on social media platforms. Along with Facebook, Twitter has been credited as
essential to the Arab Spring---both for protesters (as a means to coordinate
with one another and to communicate with the world) and for the government
response (to monitor and track protestors to crack down on dissent)
(Comninos, 2011). The events of Arab Spring have been thoroughly
discussed within the political science literature
(e.g., Lutterbeck, 2013; Koehler, 2017), but to our knowledge,
the direct \emph{coupling} between Syrian Twitter cohesion and international action
coordination has not been investigated.

\hypertarget{social-cohesion-and-social-mobilization}{%
\subsection{Social Cohesion and Social Mobilization}\label{social-cohesion-and-social-mobilization}}

@@

Just as real-world social cohesion can be
quantified, so can online cohesion. While this can be quantified in many ways,
one kind of cohesion can be found in the similarity in language use on social media.
This implicit or explicit similarity in language use can indicate shared
beliefs, thoughts, and feelings, as has been studied in research on language style
in psycholinguistics and social psychology
(Brennan \& Clark, 1996; Gonzales, Hancock, \& Pennebaker, 2010; Tausczik \& Pennebaker, 2010);
although previous work on language style
has often focused on individuals and small groups, this can be scaled up to examine
massive social-level changes (Fusaroli et al., 2015; Michel et al., 2011).
This kind of widespread online social cohesion can lead to real-world
social change.

Combining these separate threads, it is reasonable to expect that real-world actions
and online social interactions would be enmeshed, driving and being driven by
one another. Rather than expecting that only real-world social cohesion would be
driven by real-world action, we would expect interaction between real-world and
virtual dynamics: In this case, exposure to real-world violence could result in social
cohesion in the virtual world. Increased social cohesion is thus potentially
an indicator and a result of increased social organization and mobilization
on the ground.

Social media platforms provide one way to enable this kind of online social
cohesion by offering distributed information-gathering and real-time information
dissemination. Given its accessibility and proficiency as an information
exchange tool, Twitter has provided individuals involved in social movements
an opportunity to complement or even replace real-world social mobilization.
For example, during the 2012 Italian protests for global economic change, the riots in
Rome were mirrored by Twitter activity both on the day of the protests and in the
days following them (Vicari, 2013). Tweeting was
determined to be more effective locally in discussing real-world, real-time
events than traditional mainstream media platforms. As more and more
individuals rely on social media to obtain local news and information, it
becomes increasingly important to understand just how these micro-blogs
reflect real-world action.

@@

Social media has become a primary platform for social movements and
activism across the world, but the extent to which virtual activism
reflects and drives real-world mobilization is still largely unknown.
In the current work, we
conceptualize the physical and virtual movements as a dynamical system,
both constantly interacting and influencing the other. We aim to study
this by analyzing the real-world events and online social cohesion
during nearly three months of the Arab Spring. Both real- and
virtual-world collective coordination were powerful tools in the events
of the Arab Spring and are undeniably linked. This study aims to explore
the dynamics of Syrian Twitter users, who used the social media platform to raise
awareness about the Arab Spring and bring about mass mobilization.

@@

\begin{align}
\text{Two tweets similarity =}\frac{\text{inverse frequency weighting of overlap words}}{\text{inverse frequency weighting of all words}}
\end{align}

@@

Using the real deciles of social cohesion, count of all events,
count of positive events, and count of negative events, we conducted a series
of permutation tests
(one for the comparison between social cohesion and each type of event count)
to determine statistical difference from chance for each CRQA metric.
Using the \texttt{sample} function in base R (R Core Team, 2019), \(1000\) permuted time series were
created for each of the real time series.\\
CRQA was then run \(1000\) times for social cohesion and count of all
events, social cohesion and count of positive events, and social cohesion and
count of negative events using the permuted time series. This resulted in
\(1000\) results for each of the CRQA metrics for each combination of time series.
Statistical significance from chance was set at the 95th and 99th percentiles
of the metrics from the permuted CRQ analyses. Given that a permutation test
is similar to bootstrapping but does not include replacement, all points were
included in each CRQA analysis, leaving RR constant across all permutations. Thus,
RR is not included in the inferential tests.

@@

Conducting WCRQA results in CRQA metrics (e.g., RR, DET, ENTR) for each
respective window. For these analyses, we decided to focus on recurrence rate
and determinism.

@@

A permutation test for the CRQA of social cohesion and
count of positive events found no statistically significant difference for any
of the metrics other than TT (Figure \ref{fig:plot-rp-targ-pos};
\autoref{table-1}).

A permutation test for the CRQA of social cohesion and
count of negative events found a statistically significant difference for DET,
NRLINE, L, ENTR, LAM, and TT. All other metrics did not reach significance
(Figure \ref{fig:plot-rp-targ-neg}; \autoref{table-1}).

@@

In the plots (see Figures \ref{fig:plot-wcrqa-targ-all} and
\ref{fig:plot-wcrqa-source-targ-all}), the 95th percentiles are represented
by horizontal red lines, and the 99th percentiles are
represented by horizontal orange lines. The trend lines are plotted in blue.

@@

For social cohesion and count of all events, there was a consistent period
of above-chance RR from the 7th-12th and 36th-47th window. DET was largely
above chance from the 46th to the 61st window
(Figure \ref{fig:plot-wcrqa-targ-all}).

>>>>

\abstract{
Social mobilization requires effective collective coordination. In recent
years, individuals involved in social movements have taken to social media
platforms to reach broader audiences at rapid speeds. Effective virtual
mobilization similarly requires social cohesion and coordination, yet the
degree to which virtual mobilization reflects or drives real-world movements
is unclear. Here, we explore the dynamics of real-world events and social
cohesion on Twitter in Syria during the Arab Spring. Using the nonlinear
methods cross-recurrence quantification analysis and windowed cross-recurrence
quantification analysis, we investigate if counts of all daily events, daily
positive events, and daily negative events are coupled with social cohesion
found in Syrian tweets. Results indicate that virtual social cohesion is
coupled with the counts of all and negative events each day but shows a more
nuanced connection to positive events, only visible when examining finer-grained
data. This suggests that virtual mobilization does reflect
real-world movements, with the strongest connections to the frequency of all
events and the frequency of negative events. We
conclude with a discussion of implications and applications of nonlinear
methods in political science research.
}

@@

Prior to the advent of social media, people coordinated their protest behavior
using phone tree calling sheets and mimeographed flyers,
through in-person meetings of social organizations and churches, and through
word of mouth. They gathered together and hand-made protest signs that reflected
their agendas. As McAdam found, people were more likely to join movements when
they had a personal connection (McAdam, 1986; McAdam \& Paulsen, 1993):
Historically, strong ties have been
strong predictors of involvement in high-risk protest activity.

The cohesiveness and strength of social movements in the pre-Internet era could be
operationalized by the number of people turning out to protest and through the
political events that precipitated and resulted from social gatherings \emph{en masse}.
The content of placards and protest
signs provided the protest narrative in their own words. Cohesive
social movements were reflected in their cohesive placards and posters. For
example, the \enquote{I am a man} posters held by the men involved in the 1968
sanitation workers' strike reflected their internal cohesion and unified,
goal-oriented movement (Windsor, 2020).

However, the rise of the Internet and its decentralized online networks
has upended theories about recruitment to---and participation in---social
movements. Strong interpersonal ties have yielded to weak impersonal ties that
strongly predict involvement in social movements (Steinert-Threlkeld, 2017). Research
methodologies for measuring social movements have similarly evolved from
ethnographic interviews to \enquote{big data} analyses of language on social media.

Today, social media replicates and amplifies
this phenomenon during moments of social unrest. Amid the cacophony of messages
during political upheaval---such as the Arab Spring---consistent themes emerge
through hashtags, retweets, and the repeated mentions of concepts such as
regime, security, demonstration, and human rights. The cohesiveness of these
themes (i.e., whether movement participants are metaphorically singing in one voice)
reflects the real-world events happening contemporaneously with collective action in
social movements or civil conflicts.

In the present research, we investigate the relationship between linguistic
cohesion and social cohesion in times of social conflict and unrest. We tackle a
question related to the one posed by Zeitzoff (2017): How does
social media reflect on-the-ground realities during times of social conflict?
To do so, we examine patterns of linguistic cohesion on Twitter and their
connection to real-world events during
the Arab Spring uprising in Syria between March and June 2012.

Disentangling the complicated relationship between mass online mobilization---and
social mobilization in person---continues to be relevant, even as our modes
of communication change. The
political processes that transpired during and after the Arab Spring are still
in the process of being understood. As Zeitzoff (2017) recognizes,
social media is changing conflict, but the mechanisms and relationships by
which it does so are still being understood.

\hypertarget{syria-in-context}{%
\subsection{Syria in Context}\label{syria-in-context}}

The decades preceding the Arab Spring in Syria were marked by
social repression and fear. In 1971, Hafez al-Assad overthrew the previous
leader and remained in power for thirty years, presiding during
regional contentiousness and cooperation until his death in 2000. His son,
Bashar al-Assad continued the regime of authoritarian leadership as he assumed
his office. The Assad lineage hails from a minority Islamic sect (the Alawites)
in a multi-religious and multi-ethnic country comprising Christians, Sunni and
Shia Muslims, Kurds, and Palestinian refugees.

@@

Following a wave of democratization efforts across the Middle East and North
Africa, modest Syrian protests began in March 2011. The regime responded
harshly with imprisonments, detentions, censorship, and military operations as
the resistance to Assad's regime grew and spread throughout the country. The
fighting escalated dramatically in March 2012, after increased unrest in the
western city of Homs. International actors---including the Red Cross/Red Crescent
Society, the United Nations General Assembly, the Security Council, and
individual world leaders---worked toward cease-fires, negotiations,
and concessions without much success. While many high-level military and
diplomatic officials defected from the regime, numerous loyal forces
continued the regime's repression, and the Syrian people turned to social
media to help coordinate their activism and communicate to the world. Between March and
June 2012, more than 8 million tweets were sent within Syria, revealing the
details of the conflict suppressed by the government.

Syria is a multi-ethnic, multi-lingual, and multi-religious state. Kurdish
people make up between 10-15\% of the population, Alawites between 8-15\%,
Christians about 10\%. There are also much smaller sects (e.g., Ismailis,
Druze), but Sunni Muslims comprise the vast majority of society. Most people
speak Arabic as their primary language. Most policy reports addressing the current
conflict report on the cohesion of the ruling Alawite military elite, rather
than on the cohesiveness of the opposition movement. However, the Sunni
majority's disaffection with the Syrian government and their lack of political
representation provide sufficient grievance and collective group identification
to make them a cohesive force.

\hypertarget{social-media-and-social-movements}{%
\subsection{Social Media and Social Movements}\label{social-media-and-social-movements}}

Massive online social cohesion can have real-world social
manifestations. Social media platforms provide one way to enable this kind of
online social cohesion by offering distributed information-gathering and
real-time information dissemination. Given its accessibility and proficiency as
an information exchange tool, Twitter has provided individuals involved in
social movements an opportunity to complement or even replace real-world social
mobilization. For example, during the 2012 Italian protests for global economic
change, the riots in Rome were mirrored by Twitter activity both on the day of
the protests and in the days following them (Vicari, 2013). Tweeting was
determined to be more effective locally in discussing real-world, real-time
events than traditional mainstream media platforms. As more and more individuals
rely on social media to obtain local news and information, it becomes
increasingly important to understand just how these micro-blogs reflect
real-world action.

During the Arab Spring, social media provided real-time information about the
events transpiring in countries across the Middle East and North Africa. Our
understanding about the role of social media in social movements is still
evolving. Some argue that these mass movements would not have been possible
without efficient large-scale coordination, both physically on the ground and
virtually on social media platforms.

Along with Facebook, Twitter has been
credited as essential to the Arab Spring---both for protesters (as a means to
coordinate with one another and to communicate with the world) and for the
government response (to monitor and track protestors to crack down on dissent)
(Comninos, 2011). The events of Arab Spring have been thoroughly discussed
within the political science literature (e.g., Lutterbeck, 2013; Koehler, 2017), but to our knowledge, the direct \emph{coupling} between
Syrian Twitter cohesion and international action coordination has not been
investigated. In fact, some have been skeptical about the effect of social media on
collective action at all (Seay, 2014).

This skepticism can be summarized in a quote by Lisa Anderson:

\begin{quote}
The important story about the 2011
Arab revolts in Tunisia, Egypt, and Libya is not how the globalization of the
norms of civic engagement shaped the protesters' aspirations. Nor is it about
how activists used technology to share ideas and tactics. Instead, the critical
issue is how and why these ambitions and techniques resonated in their various
local contexts. (Anderson, 2011)
\end{quote}

Others have similarly expressed hesitation about the role of online
organizing. Some have argued that real-life strong ties outweigh online weak ties and
that the power of the strong ties through in-person contact is critical
to recruiting people for higher-risk activities
(Gladwell, 2010; McAdam, 1986).

Interestingly, these views may
actually help to make the case for bridging the gap between social media and
social mobilization. Such views suggest the strength of social ties in an online
environment is inextricably linked to the \emph{context} and \emph{ambitions} of the actors.
Moreover, research suggests that weak ties still retain considerable political
power: While strong ties tend to bind smaller groups, weak ties
tend to bind groups across distances (Granovetter, 1973).

Social media generates critical information about events, including whether and how
a demonstration will take place, how many
people participate, and how those in power (e.g., state, military, security apparatus)
respond. Similarly, these media platforms alert us to more global shifts within
a conflict, such as when fighting moves to a new city or region, the number of
casualties and wounded, and the types of weaponry used by parties to the
conflict (Zeitzoff, 2017). Social media is not entirely event-driven, however.
In monitoring social media channels, government forces may respond to the
perception that citizens are gaining strength and sympathy. Governments have
been shown to be quite tolerant of dissent so long as it does not metamorphose
into collective action (King, Pan, \& Roberts, 2013).

Citizens communicate with each other with increased transparency due to the
ubiquitous availability of social media such as low-cost and egalitarian
platforms like Twitter. Citizens are able to access these tools via mobile
phones and computers with Internet connectivity, providing a stream of real-time
shared information. Citizens can share the location of events like protests and military
incursions, the degree of state repression, and the number of casualties. In
turn, the government---interested in remaining in power---monitors social media
platforms to ascertain emerging threats (Comninos, 2011),
such as where new groups are forming
and where existing groups are strengthening, where gatherings are held, and who
the idea entrepreneurs are. An example of a leader
expressing his interest in
such public conversation is given by Bashar al-Assad in a speech on June 3, 2012:

\begin{quote}
I know that I have been away from the media for a long time, and I have missed
having direct contact with the citizens, but I have always been following up
with the daily occurrences and gathering the information so that my speech can
be built on what is said by the street.
\end{quote}

With few exceptions, Assad chose to keep social media platforms up and running.
Gohdes (2015) notes that regimes can attempt to thwart the growth of social
protest online and in the streets by shutting down the Internet.
During these times, however, she finds that military repression increases.
However, government information about the citizens decreases; as such, many
regimes have elected to maintain Internet connectivity as a means of monitoring
and surveilling their citizens.

At the same time, the insulation
of online social platforms allows researchers to detail real-world social dynamics
at unprecedented resolution and scale (Paxton \& Griffiths, 2017). The digital traces left by
individuals and collectively as groups are regarded as a valuable source of
information that can offer powerful insights into states of social
disequilibrium, uprisings, and escalations into civil wars.
By examining these conversations, researchers can study the degree of cohesion between
speakers and the alignment between that cohesion and their collective political
behavior.

\hypertarget{the-role-of-cohesion-in-social-movements-and-civil-wars}{%
\subsection{The Role of Cohesion in Social Movements and Civil Wars}\label{the-role-of-cohesion-in-social-movements-and-civil-wars}}

People join social and political resistance movements because they feel a sense
of connection to the cause and because they have been sufficiently convinced to
participate. Given that people are generally risk-averse, overcoming inertia
requires a strong connection to the group identity or a strong aversion to an individual,
event, or political injustice (Goldstone, 1980, 2011).
Political resistance offers participants the opportunity to \enquote{lose themselves} and
relinquish their individual identity in favor of identification with a larger
group identity, contributing to a cohesive group ready and mobilized for
action (Hoffer, 2011).

Previous research on social movements supports the idea that participants are
likely to join movements when they know and have a personal connection to
someone else already involved (McAdam, 1986). The type of collective
contentious behavior that manifests in society is a function of the available
political opportunity structures. The political opportunity structure are
\enquote{comprised \emph{(sic)} of specific configuration of resources, institutional
arrangements and historical precedents for social mobilization, which facilitate
the development of protest movements in some instances and constrain them in
others} ({\textbf{???}}). What emerges as strategies and tactics in contentious
politics are, in essence, a function of what the system makes possible. This
includes the number of people and their disposition (i.e.~nonviolent or
violent), as well as their internal structure. The more cohesive a group is, the
greater threat they pose to the establishment.

Cohesion is also a familiar concept in the study of civil conflicts, as it
pertains to group affiliation and the ability to overcome collective action
problems. Individuals must act in concert---that is, cohesively---to
accomplish the task of challenging the authority of the state (Eisinger, 1973; Francisco, 1995; Olson, 1965; Tilly, 2004).
Recent work has considered rebel recruitment for participation in
civil conflicts as a function of internal cohesion (Gates, 2002):
Rebel movement success is influenced by the degree of group cohesion, which
itself is a function of geographic distance, ideology, and ethnicity.
Within Syria, multiple rebel groups within Syria exist, sourcing fighters locally
from the disaffected Syrian population and internationally through
fighters from transnational terrorist organizations. There is considerable
intergroup competition for participation in the rebel movement and
punishment for defectors for nonparticipation (Gates, 2002).
Foreign fighters would appear to be amply motivated to fight, given their
decision to relocate from a home country that may be located in the region
or much farther away.

Another factor influencing social cohesion during social movements and
civil wars is ethnic identity and the degree of ethnic conflict associated with
the movement. Greater degrees of ethnic conflict increase the salience of
ethnic identity, leading people to identify more strongly with their own
ethnic identity (Sambanis \& Schulhofer-Wohl, 2009). As ethnic conflict
intensifies, groups seek to become more quintessentially like the typical
member of their ethnic group. This can manifest in a number of ways, including
in behavior on social media. However, this may also include a general identification
across ethnic groups in a membership of an opposition to an oppressive leader.

The civil war literature has further explored the role of social cohesion in
detail. Social homogeneity increases the likelihood of social conflict,
especially when the society is polarized (Collier \& Hoeffler, 2004); this is partly
due to nationalist groups' strong identification with a collective identity
(Cederman \& Girardin, 2007) and their social solidarity (Sambanis \& Schulhofer-Wohl, 2009). These
same processes can be similarly important for rebels (Anyanwu, 2002).
Referencing Gurr's (2000) work on minorities at
risk, Regan and Norton (2005) note that

\begin{quote}
a religious or linguistic minority
might suffer disproportionately in a given society, and this form of
aggrievement can lead to unrest across the social lines that distinguish the
minority group. Therefore, while collective cohesion is a function of targeted
maldistributions, it is individual penury that allows for this collective
mobilization. (Regan \& Norton, 2005)
\end{quote}

The body of previous work concurs that group cohesion matters to group
mobilization. We build on this by providing a quantitative approach to
understanding cohesion from multiple levels and by incorporating analytic tools
that allow us to investigate the interconnectedness of those levels.

\hypertarget{cohesion-a-multi-leveled-multi-layered-concept}{%
\subsection{Cohesion: A Multi-Leveled, Multi-Layered Concept}\label{cohesion-a-multi-leveled-multi-layered-concept}}

We contextualize linguistic cohesion with social mobilization and response to the relative
intensity of the conflict. Cohesion is a multi-layered concept referring to the
connectedness of speakers, actions, and societies. Cohesion reflects the
continuity of language and expression, and it shows the connectivity of
participants in a social movement. To address the relationship between language
and discourse and social movements, we adopt an interdisciplinary approach that
incorporates linguistics, psychology, and political science using nonlinear
dynamical systems methodology to model the relationship between language and
action.

\hypertarget{lexical-and-interlocutor-cohesion}{%
\subsubsection{Lexical and Interlocutor Cohesion}\label{lexical-and-interlocutor-cohesion}}

Just as real-world social cohesion can be quantified, so can online cohesion.
One kind of cohesion can be found in
the similarity in language use on social media. This implicit or explicit
similarity in language use can indicate shared beliefs, thoughts, and feelings,
as has been studied in research on language style in psycholinguistics and
social psychology (Brennan \& Clark, 1996; Gonzales, Hancock, \& Pennebaker, 2010; Tausczik \& Pennebaker, 2010). Although previous work on language style has often
focused on individuals and small groups, this can be scaled up to examine
massive social-level changes (Fusaroli et al., 2015; Michel et al., 2011).

The words that comprise social media posts are the first level of cohesion that
we will consider. In Figure \ref{fig:cohesion-in-tweets}, we see four
sequential tweets, each containing four words.
Each sequential tweet contains some content included in the previous tweets. The
linguistic construct of content word overlap measures the relationship across
units of meaning---here, tweets---where identical and semantically similar words
are repeated (Graesser, McNamara, Louwerse, \& Cai, 2004). These repeated lexical
items constitute a theme; on Twitter, this might be a particular hashtag, a
name, a location, or other nominal category. This localized cohesion is
referential; words are repeated and co-referred across tweets, creating a sense
of continuity and cohesiveness of meaning and message.

Typically, these analyses focus on \emph{content words} (e.g., nouns, adjectives,
verbs), which convey more meaningful semantic information than \emph{function words}
(e.g., articles, conjunctions). While function words can provide important
information about complexity and language structure
(Tausczik \& Pennebaker, 2010), content words have more specialized meaning and
are therefore more useful when researchers want to understand \emph{what} people are
discussing, rather than \emph{how} they are discussing it (Graesser et al., 2004; McNamara, Graesser, McCarthy, \& Cai, 2014).

In estimating cohesion in Twitter posts, tweets that share content words
will yield higher cohesion values, even among speakers with opposing
views on the subject matter. Repetition of terms and concepts is captured by
reiteration, which is a type of lexical cohesion in which a subsequent item
repeats a previous item, or refers to it through synonymous language
(Halliday \& Hasan, 1976). In social media, we can conceive of retweets (i.e., reposts
or forwards of previous messages) as a form of reiteration
that increase message cohesion. Since retweets replicate verbatim some part of
another tweet, these tweets would also have higher cohesion. Hashtags function
similarly and are included in our analysis.

\begin{figure}
\includegraphics[width=1\linewidth]{./figs/cohesion-in-tweets} \caption{A visual representation of cohesion in tweets. Content word overlap indicates a shared relationship among tweets. The shared words--either identical or semantically similar--constitute a theme around which cohesion is occuring.}\label{fig:cohesion-in-tweets}
\end{figure}

The use of similar language can provide insight into \emph{interlocutor cohesion}, or
the degree to which people are talking about similar things. This can---as noted
earlier---include people who are talking about similar things from radically
perspectives, as well as those who are talking about something from the same
perspective. Previous research from political science has investigated these
ideas (Windsor, 2020): The \enquote{Occupy Wall Street} movement has been held up as an
example of a movement with low interlocutor and linguistic cohesion because of
its wide range of signs and slogans, while the \enquote{I am a man} protest (captured by
Ernest Withers during the 1968 sanitation workers' strike in Memphis, TN) or the
\enquote{Black Lives Matter} placards (following the uprising in Ferguson, MO) have high
lexical and interlocutor cohesion (Edgar \& Johnson, 2018).

Fed by lexical cohesion, interlocutor cohesion ebbs and flows along a timeline
contextualized by variation in real world events. Interlocutor cohesion may
increase preceding an in-person social mobilization event, alerting a repressive
government that collective mobilization is happening. Interlocutor cohesion may
also respond to exogenous events such as spontaneous, unprovoked militarized
conflict. If government forces target a new location using military weapons,
this can precipitate online mobilization and discussion of the event.

Given the importance of lexical and interlocutor cohesion for messaging in
political movements, lexical cohesion may be an important bit of trace data for
researchers interested in leveraging social media data to analyze social
movements. Similar to large-scale studies of sociocultural movements (sometimes
referred to as \emph{culturomics}; e.g., Michel et al., 2011), political
scientists can track the ways in which conversations on social media coalesce or
diffuse around similar content words. In this way, it matters relatively little
whether the interlocutors have precisely the same views: The point is to
understand whether the society-level conversation is converging on similar
language.

\hypertarget{social-and-political-cohesion}{%
\subsubsection{Social and Political Cohesion}\label{social-and-political-cohesion}}

@@

At the same time, behavioral similarity in certain contexts has been
connected to important psychological and social effects. For example, shared
language is associated with rapport (both in face-to-face and computer-mediated
conversations; Niederhoffer \& Pennebaker, 2002; Riordan, Dale, Kreuz, \& Olney, 2011) and
can even be connected to better performance on shared tasks (as long as
the language is relevant to the task at hand; Fusaroli et al., 2012). Shared
movement is associated with greater group bonding and greater willingness
to contribute to the group (Wiltermuth \& Heath, 2009), and shared identity
is associated with a greater willingness to adopt the group's goals and
motivations (Walton, Cohen, Cwir, \& Spencer, 2012).

We should expect to see increased linguistic cohesion among people participating
in online discussions about an idea, event, individual, or movement. Given that
Syrian protesters are known to have coordinated their activity through Twitter
by discussing their own events (e.g., planned protests, active demonstrations)
and their reactions to events happening around them (e.g., government reactions,
international action or inaction), the cohesion in social media should be
expected to correlate with the social cohesion of the group (Zeitzoff, 2011).

This social cohesion can have real and wide-ranging impacts on a political
movement and the people in it. A socially cohesive group is inherently more
threatening to the existing regime because it reflects collective opposition
momentum and the professionalization of potential alternate leadership. Groups
with high social cohesion are more difficult to divide and conquer, and more
cohesive messages may also garner increased international support. Global
insurgency movements exhibit similar patterns in their social structure
(Bohorquez, Gourley, Dixon, Spagat, \& Johnson, 2009). Each violent group has a stable state. Exogenous events
can create perturbations that push the group to become either more cohesive or
more fragmented. Rather than frenetic, patchwork messages via Twitter, emergent
or intentional cohesive online messaging---using similar language, paralleling
in-person messaging (Windsor, 2020)---are a sign of increased organizational
structure, inciting a government response.

Thus, lexical and interpersonal cohesion are hypothesized to be tied to social
and political cohesion. Shared language in an online environment may lead to the
development or expression of a shared identity (Wiltermuth \& Heath, 2009). This,
in turn, may drive individuals' propensity to engage in similar coordinated
behavior on behalf of the group (Walton et al., 2012; Wiltermuth \& Heath, 2009).
Taken together, these two processes may grow movement participants' weak and
strong ties (Gladwell, 2010; Granovetter, 1973; McAdam, 1986), strengthening the movement overall by expanding and
deepening heterogenous ties. Therefore,
in the context of social conflict, the linguistic cohesion
of the online conversation should correspond to the degree of
social cohesion on the ground.

Steinert-Threlkeld (2017) modeled a similar relationship
between social media and social mobilization alongside event data. This
groundbreaking work demonstrates that participants on the fringe of a social
movement contribute substantially to the overall level of protest. In addition
to providing new methods to this growing field of study, we build on this
previous work by demonstrating how sociolinguistic cohesion evolves in
response to negative and positive exogenous events.

Political and social cohesion in the world can be modeled using event data which
encodes who does what to whom on a given day. This could include events such as
the government imposing a blockade or fighting with artillery or tanks on
a civilian group. CAMEO (Conflict and Mediation Event Observations) encode
a set of activities with numerical values, rescaled to reflect a
conflict-cooperation scale (Goldstein, 1992; Schrodt \& Hall, 2006). These event
codes include a range of behaviors, from making public statements and appeals,
to demanding, threatening, protesting, and using unconventional mass violence.

\hypertarget{sociolinguistic-cohesion-connectedness-of-speakers-ideas-and-events}{%
\subsubsection{Sociolinguistic Cohesion: Connectedness of Speakers, Ideas, and Events}\label{sociolinguistic-cohesion-connectedness-of-speakers-ideas-and-events}}

We tie the constructs of lexical, interlocutor, and social cohesion together in
a theoretical framework called \emph{sociolinguistic cohesion}, as shown in Figure
\ref{fig:theoretical-framework}. Several interconnected mechanisms are in
motion here. Previous values of linguistic cohesion should influence future
values of linguistic cohesion and future actions in the real world space.
Previous events influence future events and future linguistic cohesion in
social media.

Sociolinguistic cohesion is not a static concept. Just as social movements
experience increased participation and setbacks---indicating dynamic
relationships among participants and between participants and the sociopolitical
environment---sociolinguistic cohesion is also dynamic, reflecting the changing
states of the political sphere. Other scholars have noted the diffusion of
messages whereby they reach a critical mass, whereby \enquote{behaviors, beliefs, and
preferences} spread within a population (Barash, Cameron, \& Macy, 2012). Cohesion
changes over time in response to external events and the internal dynamics of
the social movement. This can occur even if the social interaction occurs entirely
online (Centola \& Macy, 2007).

Related previous work has also explored diffusion of hashtags in viral contagion in
Nigeria, a phenomenon similar to what we identify as sociolinguistic cohesion
(Fink, Schmidt, Barash, Cameron, \& Macy, 2016). The prevalence and rate of particular hashtags were
contributed to the overall sociolinguistic cohesion. The more that people
reuse hashtags or common phrases, the greater the sociolinguistic cohesion when
this phenomenon occurs during socially significant times like protests
(State \& Adamic, 2015).

Combining these separate threads, it is reasonable to expect that real-world
actions and online social interactions would be enmeshed, driving and being
driven by one another. Rather than expecting that only real-world social
cohesion would be driven by real-world action, we would expect interaction
between real-world and online dynamics: In this case, exposure to real-world
violence could result in social cohesion in the virtual world. Increased social
cohesion is thus potentially an indicator and a result of increased social
organization and mobilization on the ground. However, this occurs within a context
that is affected by local and international events; both
exogenous events (like intervention by another country or
international organization like the United Nations) and
endogenous events (like the recruitment of new members to the rebel or
resistance movement) can affect cohesion.

From a dynamical systems perspective, online social cohesion and
in-person action can be considered two complementary
components of a single system, mutually constrained and constantly interacting
and influencing one another (Mahner \& Bunge, 1997). Investigation of the
coupling of real-world mobilization and social cohesion could provide new
insight into the power of social media during mobilization and the ways in which
individuals utilize it to bring about, or respond to, social change.
Increasingly, social media platforms are being utilized during social
mobilization (Koopmans, 2004; McAdam, Tarrow, \& Tilly, 2003),
but the relationship between virtual and real-world mobilization
is still largely unknown. The present work aims to understand how physical and
virtual movements coexist and interact from a dynamical systems perspective.

Lexical cohesion, then, is be a useful indicator of social cohesion in
political crisis. Lexical cohesion happens at the word level. Interlocutor
cohesion happens between speakers. Political cohesion happens at the social
level during moments of political crisis, like civil wars and social movements.
Sociolinguistic cohesion is comprised of lexical, interlocutor, and
socio-political events, where the words and their context reveal the disposition
of socially connected linguistic communities. Sociolinguistic cohesion combines
the elements of lexical cohesion in that it captures the degree of similarity
between what is said, and how it is said. It accounts for the cohesion between
multiple speakers in an environment whose language evolves to become more
aligned and similar. Sociolinguistic cohesion also addresses how groups involved
in political rebellion must have internal cohesion to overcome the collective
action problem to mount sustained resistance to state authority.

\begin{figure}
\includegraphics[width=1\linewidth]{./figs/theoretical-framework} \caption{The theoretical framework for sociolinguistic cohesion. Real-world events and linguistic cohesion continually interact and influence one another throughout time.}\label{fig:theoretical-framework}
\end{figure}

@@

Social media has become a primary platform for social movements and activism
across the world, but the extent to which virtual activism reflects and drives
real-world mobilization is still largely unknown. In the current work, we
conceptualize the physical and virtual movements as a dynamical system, both
constantly interacting and influencing the other. We aim to study this by
analyzing the real-world events and online social cohesion during nearly three
months of the Arab Spring. Both real- and virtual-world collective coordination
were powerful tools in the events of the Arab Spring and are undeniably linked.
This study aims to explore the dynamics of Twitter users in Syria, who used the
social media platform to raise awareness about the Arab Spring and bring about
mass mobilization.

@@

\begin{align}
\text{Two tweets' similarity =}\frac{\text{inverse frequency weighting of overlap
words}}{\text{inverse frequency weighting of all words}}
\end{align}

@@

Using the real deciles of social cohesion, count of all events, count of
positive events, and count of negative events, we conducted a series of
permutation tests (one for the comparison between social cohesion and each type
of event count) to determine statistical difference from chance for each CRQA
metric. Using the \texttt{sample} function in base R (R Core Team, 2019), \(1000\) permuted time
series were created for each of the real time series. CRQA was then run \(1000\)
times for social cohesion and count of all events, social cohesion and count of
positive events, and social cohesion and count of negative events using the
permuted time series. This resulted in \(1000\) results for each of the CRQA
metrics for each combination of time series. Statistical significance from
chance was set at the 95th and 99th percentiles of the metrics from the permuted
CRQ analyses. Given that a permutation test is similar to bootstrapping but does
not include replacement, all points were included in each CRQA analysis, leaving
RR constant across all permutations. Thus, RR is not included in the inferential
tests.

@@

For WCRQA analyses, we decided to focus on RR and DET.

@@

A permutation test for the CRQA of social cohesion and count of positive events
found no statistically significant difference from chance for any of the metrics
other than TT (Figure \ref{fig:plot-rp-targ-pos}; \autoref{table-1}).

A permutation test for the CRQA of social cohesion and count of negative events
found a statistically significant difference from chance for DET, NRLINE, L,
ENTR, LAM, and TT. All other metrics did not reach significance (Figure
\ref{fig:plot-rp-targ-neg}; \autoref{table-1}).

@@

In the WCRQA plots (see Figures \ref{fig:plot-wcrqa-targ-all} and
\ref{fig:plot-wcrqa-source-targ-all}), the 95th percentiles are represented by
horizontal red lines, and the 99th percentiles are represented by horizontal
orange lines. The trend lines are plotted in blue.

@@

For social cohesion and count of all events, there was a consistent period of
above-chance RR from the 7th--12th and 32nd--47th window. In addition to some
one-week peaks at the beginning of the time period, DET was largely above chance
from the 46th to the 61st window, with some exceptions (Figure
\ref{fig:plot-wcrqa-targ-all}).
